---
title: "heatmap_ggplot"
author: "Cheryl Logan"
date: "4/13/2018"
output: html_document
---
R code for making heatmaps in ggplot2

Install necessary packages and load libraries. You only need to install the above packages once, but need to load the libraries every time you open up a new R session.

```{r, eval = FALSE}
library(tidyverse) # after install.packages('tidyverse')
library(reshape2)  # after install.packages('reshape2')
library(ggplot2)   # after install.packages('ggplot2')
```

Display working directory (equivalent of pwd in terminal)

```{r, eval = FALSE}
getwd()
```

If necessary, change the working directory to location of your expression data file
```{r, eval = FALSE}
setwd("/Users/github/bio430")
```

Load expression data (log 2 centered)
```{r, eval = FALSE}
data <- read_csv("diffExpr.P0.01_C1.matrix.log2.centered.csv")
head(data)
```

Reshape data into 3 columns using `gather`

```{r, eval = FALSE}
df_heatmap <- gather(data, "treatment","expr", 2:21)
head(df_heatmap)
```
Make heatmap

```{r, eval = FALSE}
ggplot(df_heatmap, aes(x=treatment, y=ContigName)) +
  geom_tile(aes(fill = expr)) +
  scale_fill_gradient2(low="green", mid="black", high="red", midpoint=0, limits=c(-2,2)) +
  ylab("Genes ") +
  xlab("Treatments") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "Log2 Fold Change")
```

For additional info about how to modify this input:

http://databiomics.com/how-to-create-a-fast-and-easy-heatmap-with-ggplot2/